Convex Surrogates via Polyhedral Losses

### 2019-01-30 cvx-fin.tex
proof of polyhedral implies embed

\begin{theorem}\label{thm:polyhedral-embed}
  Every polyhedral loss embeds a finite elicitable property.
  Moreover, a polyhedral loss indirectly elicits a finite elicitable property $\gamma$ if and only if it embeds a property which refines $\gamma$.
\end{theorem}
\begin{proof}
  \raf{Turns out I don't have a proof.  Seems heavily related to the separated link stuff...}
\end{proof}

@@

\begin{lemma}\label{lem:unique-opt-on-inter}
  Let $\Gamma:\simplex \toto \R'$ be an elicitable property such that $\trim(\Gamma)$ is finite.
  For all $u \in \R'$, if $p \in \inter{\Gamma_u}$, then $\Gamma(p) = \{u\}$.
\end{lemma}
\begin{proof}
\jessiet{Come back and formalize}
  We'll use the relationship between nonredundant properties and subgradients of the expected score $G$.
  Now consider the following convex analysis fact: for any convex function and any two point which share a subgradient, the function is flat between those points.
  In particular, $G$ is flat on $\Gamma_u$, and therefore $G$ must be differentiable on the interior of $\Gamma_u$.
  Therefore, $G$ must only have one subgradient on $\inter{\Gamma_u}$.
  \bigskip
  \jessie{Formalization below}
  \hline
  Consider that $\trim(\Gamma)$ is nonredundant, and $p \in \inter{\Gamma_u}$ only if $\Gamma_u \in \trim(\Gamma)$ by Lemma~\ref{lem:trim-full-dim}, as a closed, convex set is full-dimensional if and only if it has nonempty interior. \jessiet{Sanity check here?}
  Since $\Gamma$ is elicitable, the score $G$ must be convex.
  For $w \in \R'$, if $\Gamma_w \subseteq \Gamma_u$, then we contradict the nonredundancy of $\trim(\Gamma)$, so we only worry about the case where $\Gamma_w \not \subseteq \Gamma_u$..

@@

  Let's say there is a $q \in \Gamma_w \cap \inter{\Gamma_u}$ and $\q' \in \Gamma_w \cap \inter{\Gamma_v}$.
  Additionally, $\conv(\{q,q'\}) \subseteq \Gamma_w$, and observe that $\conv(\{q,q'\})$ is not countable.
  Since there is a bijection from $\Gamma(p)$ to $\partial_p G(p)$, we then note that $|\partial_pG(p)| > 1$ on an uncountable set, and $\partial_p G(p)$ is not constant on $\conv(\{q,q'\})$.
  This contradicts the convexity of $G$, and in turn, the elicitability of $\Gamma$ and $\trim(\Gamma)$.

@@

\begin{lemma}
  Let $f_1,\ldots,f_k:\reals^d\to\reals$ be polyhedral convex functions.
  The power diagram induced by $\sum_{i=1}^k p_i f_i$ is the same for all $p \in \inter\simplex$.
\end{lemma}
\begin{proof}
  For any convex function $g$ with epigraph $P$, the proof of~\citet[Theorem 4]{aurenhammer1987power} shows that the power diagram induced by $g$ is determined by the facets of $P$.
  Let $F$ be a facet of $P$, and $F'$ its projection down to $\reals^d$.
  It follows that $g|_{F'}$ is affine, and thus $g$ is differentiable on $\inter F'$ with constant derivative $d\in\reals^d$.
  Conversely, for any subgradient $d'$ of $g$, the set of points $\{x\in\reals^d : d'\in\partial g(x)\}$ is the projection of a face of $P$; we conclude that $F = \{(x,g(x))\in\reals^{d+1} : d\in\partial g(x)\}$ and $F' = \{x\in\reals^d : d\in\partial g(x)\}$.

>>>>

\begin{theorem}\label{thm:polyhedral-embed}
  Every polyhedral loss embeds a finite elicitable property.
  Moreover, a polyhedral loss indirectly elicits a finite elicitable property $\gamma$ if and only if it embeds a property which refines $\gamma$.
\end{theorem}
\begin{proof}
  \raf{This proof is kind of awesome.}

  Let $L:\reals^d\to\reals_+$ be a polyhedral loss.
  For all $p$, let $P(p)$ be the epigraph of the convex function $u\mapsto p\cdot L(u)$.
  From Lemma~\ref{lem:polyhedral-pd-same}, we have that the power diagram induced by the projection of $P(p)$ onto $\reals^d$ is constant whenever $p\in\inter\simplex$.
  Let $q\in\inter\simplex$ be the uniform distribution on $\Y$, and $V_\Y$ be the set of vertices of $P(q)$ projected onto $\reals^d$.
  By the above, this set is the same had we replaced $q$ by any $p\in\inter\simplex$.

  Now let $\Gamma := \Gamma[L]$.
  We claim that for all $p\in\inter\simplex$, if $\Gamma(p) \cap V_\Y \neq \emptyset$.
  To see this, let $u \in \Gamma(p)$, and $u' = (u,p\cdot L(u)) \in P(p)$.
  The optimality of $u$ is equivalent to $u$ being contained in the face exposed by the normal $(0,\ldots,0,-1)\in\reals^{d+1}$, which is a face of $P(p)$.
  Let $v'\in\reals^{d+1}$ be a vertex on such a face, and $v\in V_\Y$ its projection onto $\reals^d$.
  Then $v$ is also optimal, and therefore $v\in\Gamma(p)$.

  Now consider $\Y'\subset \Y$.
  Applying the above argument on distributions $p$ with support exactly $\Y'$, we have a similar guarantee: a finite set $V_{\Y'}$ such that for all $p$ with support exactly $\Y'$, we have $\Gamma(p) \cap V_{\Y'} \neq \emptyset$.
  (When $\Y' = \{y\}$ is a singleton, we simply take the projected vertices of $L(\cdot)_y$.)

  Thus, taking $V = \bigcup_{\Y'\subseteq\Y}$, we have for all $p\in\simplex$ that $\Gamma(p) \cap V \neq \emptyset$.
  This implies that $\trim(\Gamma) \subseteq \{\Gamma_v : v\in V\}$, which is finite, so Proposition~\ref{prop:embed-trim} now gives the conclusion.
\end{proof}

@@

\begin{lemma}\label{lem:unique-opt-on-inter}
  Let $\Gamma:\simplex \toto \R'$ be an elicitable property such that $\trim(\Gamma)$ is finite.
  For all $u \in \R'$, if $p \in \inter{\Gamma_u}$, then $\Gamma(p) = \{u\}$.
\end{lemma}
\begin{proof}
\jessiet{Come back and formalize}
  We'll use the relationship between nonredundant properties and subgradients of the expected score $G$.
  Now consider the following convex analysis fact: for any convex function and any two point which share a subgradient, the function is flat between those points.
  In particular, $G$ is flat on $\Gamma_u$, and therefore $G$ must be differentiable on the interior of $\Gamma_u$.
  Therefore, $G$ must only have one subgradient on $\inter{\Gamma_u}$.
  \bigskip
  \jessie{Formalization below}
  \hrule
  Consider that $\trim(\Gamma)$ is nonredundant, and $p \in \inter{\Gamma_u}$ only if $\Gamma_u \in \trim(\Gamma)$ by Lemma~\ref{lem:trim-full-dim}, as a closed, convex set is full-dimensional if and only if it has nonempty interior. \jessiet{Sanity check here?}
  Since $\Gamma$ is elicitable, the score $G$ must be convex.
  For $w \in \R'$, if $\Gamma_w \subseteq \Gamma_u$, then we contradict the nonredundancy of $\trim(\Gamma)$, so we only worry about the case where $\Gamma_w \not \subseteq \Gamma_u$..

@@

  Let's say there is a $q \in \Gamma_w \cap \inter{\Gamma_u}$ and $q' \in \Gamma_w \cap \inter{\Gamma_v}$.
  Additionally, $\conv(\{q,q'\}) \subseteq \Gamma_w$, and observe that $\conv(\{q,q'\})$ is not countable.
  Since there is a bijection from $\Gamma(p)$ to $\partial_p G(p)$, we then note that $|\partial_pG(p)| > 1$ on an uncountable set, and $\partial_p G(p)$ is not constant on $\conv(\{q,q'\})$.
  This contradicts the convexity of $G$, and in turn, the elicitability of $\Gamma$ and $\trim(\Gamma)$.

@@

\begin{lemma}\label{lem:polyhedral-pd-same}
  Let $f_1,\ldots,f_k:\reals^d\to\reals$ be polyhedral convex functions.
  The power diagram induced by $\sum_{i=1}^k p_i f_i$ is the same for all $p \in \inter\simplex$.
\end{lemma}
\begin{proof}
  For any convex function $g$ with epigraph $P$, the proof of~\citet[Theorem 4]{aurenhammer1987power} shows that the power diagram induced by $g$ is determined by the facets of $P$.
  Let $F$ be a facet of $P$, and $F'$ its projection down to $\reals^d$.
  It follows that $g|_{F'}$ is affine, and thus $g$ is differentiable on $\inter F'$ with constant derivative $d\in\reals^d$.
  Conversely, for any subgradient $d'$ of $g$, the set of points $\{x\in\reals^d : d'\in\partial g(x)\}$ is the projection of a face of $P$; we conclude that $F = \{(x,g(x))\in\reals^{d+1} : d\in\partial g(x)\}$ and $F' = \{x\in\reals^d : d\in\partial g(x)\}$.
